---
title: "Credit Risk Demo: Environment & Data Setup"
format: html
execute:
  warning: false
  message: false
---

**Run this document once** before the credit risk demo to prepare the
environment, create the database layout, and generate synthetic lending data.

After this document completes, the demo (`workspace_credit_risk_demo.qmd`)
runs from Section 1 with no setup delays.

**What this document does:**

1. Load R packages: tidymodels, XGBoost, ranger
2. Connect to Snowflake via DBI/odbc
3. Create a dedicated database with isolated schemas for source data,
   Feature Store, training artefacts, and model registry
4. Generate three synthetic lending tables (~50K loans, 10K customers, ~500K payments)
5. Validate everything is ready

## 1. Setup

::: {.callout-note}
In Positron, the R environment and packages are managed by your local R
installation. Install any missing packages with `pak::pkg_install()` or
`install.packages()` before running this document.
:::

Load packages and verify versions:

```{r}
#| label: load-libraries
library(DBI)
library(odbc)
library(yaml)
library(tidymodels)
library(xgboost)
library(ggplot2)
library(dplyr)
library(scales)
cat("tidymodels", as.character(packageVersion("tidymodels")), "loaded\n")
cat("xgboost",   as.character(packageVersion("xgboost")),   "loaded\n")
```

## 3. Connect and Create Database Layout

Connect to Snowflake and configure the database layout. Create four isolated
schemas:

| Schema | Purpose |
|---|---|
| `RAW_DATA` | Synthetic source tables (customers, payments, loans) |
| `FEATURES` | Feature Store (entity, feature views, dynamic tables) |
| `TRAINING` | Materialised training datasets |
| `MODELS` | Model Registry (versioned models, metrics) |

```{r}
#| label: connect
conn <- DBI::dbConnect(
  odbc::snowflake(),
  Database  = "CREDIT_RISK_ML",
  Schema    = "PUBLIC",
  Warehouse = "DEFAULT_WH"
)
conn
```

```{r}
#| label: config
# Read schema names from config file; fall back to defaults if not present
# cfg <- tryCatch(
#   yaml::read_yaml("snowflaker_credit_risk_config.yaml")$credit_risk,
#   error = function(e) list()
# )
cfg <- list(
  database             = Sys.getenv("CR_DB",'CREDIT_RISK_ML'),
  source_schema        = Sys.getenv("CR_SOURCE_SCHEMA",'RAW_DATA'),
  feature_store_schema = Sys.getenv("CR_FEATURE_STORE_SCHEMA",'FEATURES'),
  training_schema      = Sys.getenv("CR_TRAINING_SCHEMA",'TRAINING'),
  registry_schema      = Sys.getenv("CR_REGISTRY_SCHEMA",'MODELS')
)
cr_db       <- cfg$database             %||% "CREDIT_RISK_ML"
cr_source   <- cfg$source_schema        %||% "RAW_DATA"
cr_features <- cfg$feature_store_schema %||% "FEATURES"
cr_training <- cfg$training_schema      %||% "TRAINING"
cr_registry <- cfg$registry_schema      %||% "MODELS"

fqn_source  <- function(name) paste(cr_db, cr_source,   name, sep = ".")
fqn_feature <- function(name) paste(cr_db, cr_features, name, sep = ".")
fqn_model   <- function(name) paste(cr_db, cr_registry, name, sep = ".")

cat("Database layout:\n")
cat(sprintf("  Database:       %s\n", cr_db))
cat(sprintf("  Source data:    %s.%s\n", cr_db, cr_source))
cat(sprintf("  Feature Store:  %s.%s\n", cr_db, cr_features))
cat(sprintf("  Training:       %s.%s\n", cr_db, cr_training))
cat(sprintf("  Model Registry: %s.%s\n", cr_db, cr_registry))
```

### Create Database and Schemas

```{r}
#| label: create-db
DBI::dbExecute(conn, paste("CREATE DATABASE IF NOT EXISTS", cr_db))

for (schema in c(cr_source, cr_features, cr_training, cr_registry)) {
  DBI::dbExecute(conn, paste(
    "CREATE SCHEMA IF NOT EXISTS", paste(cr_db, schema, sep = ".")
  ))
}

cat(sprintf("Database %s ready with 4 schemas.\n", cr_db))
```

## 4. Generate Synthetic Data

Three tables simulate a lending platform's data warehouse.
All data is generated in Snowflake using SQL — no external files needed.

- **CREDIT_CUSTOMERS** — 10K customers with income, employment, housing
- **CREDIT_PAYMENTS** — ~200K payment records with delinquency info
- **CREDIT_LOANS** — ~50K applications with a ~15% default rate

The default probability is correlated with income, employment length, and
payment behaviour so our models have real signal to learn.

```{r}
#| label: table-names
tbl_customers <- fqn_source("CREDIT_CUSTOMERS")
tbl_payments  <- fqn_source("CREDIT_PAYMENTS")
tbl_loans     <- fqn_source("CREDIT_LOANS")

cat("Tables will be created as:\n")
cat(" ", tbl_customers, "\n")
cat(" ", tbl_payments,  "\n")
cat(" ", tbl_loans,     "\n")
```

### Customer Demographics

```{r}
#| label: create-customers
DBI::dbExecute(conn, paste(
  "CREATE OR REPLACE TABLE", tbl_customers, "AS",
  "SELECT",
  "  ROW_NUMBER() OVER (ORDER BY SEQ4()) AS customer_id,",
  "  GREATEST(21, LEAST(75, 18 + UNIFORM(0, 57, RANDOM()))) AS age,",
  "  GREATEST(15000,",
  "    ROUND(UNIFORM(20000, 200000, RANDOM()), -2)) AS annual_income,",
  "  GREATEST(0, LEAST(40,",
  "    ROUND(UNIFORM(0, 300, RANDOM()) / 10.0, 1))) AS employment_length_years,",
  "  ARRAY_CONSTRUCT('RENT','OWN','MORTGAGE')[",
  "    UNIFORM(0, 2, RANDOM())] :: STRING AS home_ownership,",
  "  ARRAY_CONSTRUCT(",
  "    'CA','TX','NY','FL','IL','PA','OH','GA','NC','MI',",
  "    'NJ','VA','WA','AZ','MA','CO','TN','MD','MN','IN')[",
  "    UNIFORM(0, 19, RANDOM())] :: STRING AS state",
  "FROM TABLE(GENERATOR(ROWCOUNT => 10000))"
))
cat(sprintf("Created %s\n", tbl_customers))
DBI::dbGetQuery(conn, paste("SELECT COUNT(*) AS N FROM", tbl_customers))
```

### Payment History

```{r}
#| label: create-payments
DBI::dbExecute(conn, paste(
  "CREATE OR REPLACE TABLE", tbl_payments, "AS",
  "WITH row_gen AS (",
  "  SELECT MOD(SEQ4(), 10000) + 1 AS customer_id",
  "  FROM TABLE(GENERATOR(ROWCOUNT => 200000))",
  "),",
  "payment_base AS (",
  "  SELECT",
  "    r.customer_id,",
  "    DATEADD('day', -UNIFORM(1, 1095, RANDOM()),",
  "      CURRENT_DATE()) AS payment_date,",
  "    GREATEST(50, ROUND(",
  "      c.annual_income / 12 * 0.15",
  "      + UNIFORM(-400, 400, RANDOM()), 2)) AS payment_amount,",
  "    CASE",
  "      WHEN UNIFORM(0, 100, RANDOM()) <",
  "        GREATEST(5, 40 - c.employment_length_years * 2",
  "          - c.annual_income / 20000)",
  "      THEN UNIFORM(1, 60, RANDOM())",
  "      ELSE 0",
  "    END AS days_past_due",
  "  FROM row_gen r",
  "  JOIN", tbl_customers, "c ON r.customer_id = c.customer_id",
  ")",
  "SELECT",
  "  customer_id, payment_date, payment_amount, days_past_due,",
  "  CASE",
  "    WHEN days_past_due = 0 THEN 'ON_TIME'",
  "    WHEN days_past_due <= 30 THEN 'LATE'",
  "    ELSE 'MISSED'",
  "  END AS payment_status",
  "FROM payment_base"
))
cat(sprintf("Created %s\n", tbl_payments))
DBI::dbGetQuery(conn, paste("SELECT COUNT(*) AS N FROM", tbl_payments))
```

### Loan Applications

```{r}
#| label: create-loans
DBI::dbExecute(conn, paste(
  "CREATE OR REPLACE TABLE", tbl_loans, "AS",
  "WITH row_gen AS (",
  "  SELECT",
  "    MOD(SEQ4(), 10000) + 1 AS customer_id,",
  "    ROW_NUMBER() OVER (ORDER BY SEQ4()) AS application_id",
  "  FROM TABLE(GENERATOR(ROWCOUNT => 50000))",
  "),",
  "pay_summary AS (",
  "  SELECT customer_id,",
  "    AVG(days_past_due) AS avg_dpd,",
  "    SUM(CASE WHEN payment_status = 'MISSED' THEN 1 ELSE 0 END)",
  "      / COUNT(*) AS miss_rate",
  "  FROM", tbl_payments,
  "  GROUP BY customer_id",
  "),",
  "apps AS (",
  "  SELECT",
  "    r.application_id,",
  "    r.customer_id,",
  "    DATEADD('day', -UNIFORM(1, 730, RANDOM()),",
  "      CURRENT_DATE()) AS application_date,",
  "    ROUND(UNIFORM(1000, 50000, RANDOM()), -2) AS loan_amount,",
  "    ARRAY_CONSTRUCT(12, 24, 36, 48, 60)[",
  "      UNIFORM(0, 4, RANDOM())] :: INTEGER AS loan_term_months,",
  "    ROUND(UNIFORM(3.5, 24.0, RANDOM())::FLOAT, 2) AS interest_rate,",
  "    ARRAY_CONSTRUCT(",
  "      'DEBT_CONSOLIDATION','HOME_IMPROVEMENT','BUSINESS',",
  "      'AUTO','MEDICAL','EDUCATION','OTHER')[",
  "      UNIFORM(0, 6, RANDOM())] :: STRING AS loan_purpose,",
  "    CASE",
  "      WHEN UNIFORM(0, 100, RANDOM()) <",
  "        GREATEST(2, LEAST(60,",
  "          35",
  "          - c.annual_income / 10000",
  "          - c.employment_length_years",
  "          + COALESCE(p.avg_dpd, 5) * 0.5",
  "          + COALESCE(p.miss_rate, 0.1) * 30",
  "        ))",
  "      THEN 1 ELSE 0",
  "    END AS defaulted",
  "  FROM row_gen r",
  "  JOIN", tbl_customers, "c ON r.customer_id = c.customer_id",
  "  LEFT JOIN pay_summary p ON r.customer_id = p.customer_id",
  ")",
  "SELECT * FROM apps"
))
cat(sprintf("Created %s\n", tbl_loans))

DBI::dbGetQuery(conn, paste(
  "SELECT COUNT(*) AS total_apps,",
  "  SUM(defaulted) AS defaults,",
  "  ROUND(AVG(defaulted) * 100, 1) AS default_rate_pct",
  "FROM", tbl_loans
))
```

## 5. Validation

Confirm the database, schemas, and source tables are ready for the demo.

```{r}
#| label: validation
cat("=== Setup Validation ===\n\n")

cat(sprintf("Database: %s\n", cr_db))
cat("Schemas:\n")
for (s in c(cr_source, cr_features, cr_training, cr_registry)) {
  cat(sprintf("  %s.%s\n", cr_db, s))
}
cat("\n")

cat("Tables:\n")
for (tbl in c(tbl_customers, tbl_payments, tbl_loans)) {
  n <- DBI::dbGetQuery(conn, paste("SELECT COUNT(*) AS N FROM", tbl))
  cat(sprintf("  %s: %s rows\n", tbl, format(n$N, big.mark = ",")))
}

cat("\nSetup complete. You can now run workspace_credit_risk_demo.qmd.\n")
```

## 6. Cleanup (Optional)

Uncomment to tear down everything created by this setup document.

::: {.callout-warning}
This drops the entire database including all schemas, tables, Feature Views,
and registered models.
:::

```{r}
#| label: cleanup
# DBI::dbExecute(conn, paste("DROP DATABASE IF EXISTS", cr_db))
# cat(sprintf("Dropped database %s\n", cr_db))

cat("Cleanup section -- uncomment the line above to drop everything\n")
```
