---
title: "Credit Risk Scoring: End-to-End ML with snowflakeR"
subtitle: "Feature Store → tidymodels → Model Registry → SPCS Deployment"
format: html
---

An end-to-end credit risk scoring workflow demonstrating how a fintech company
can build, evaluate, and deploy credit-risk models entirely within Snowflake
using R and the `snowflakeR` package.

**What you'll do:**

1. Explore lending data already in Snowflake
2. Build reusable features in the Snowflake Feature Store
3. Train competing models with tidymodels (logistic regression vs XGBoost)
4. Evaluate and compare model performance
5. Register both models in the Snowflake Model Registry with metrics
6. Select the champion version programmatically and deploy via SPCS
7. Score new applications via REST and SQL
8. Verify autocapture (ML Observability)

**Database layout** (created by `00_setup.qmd`):

| Schema | Contents |
|--------|----------|
| `RAW_DATA` | Source tables — customers, payments, loans |
| `FEATURES` | Feature Store (entity, feature views, dynamic tables) |
| `TRAINING` | Materialised training datasets |
| `MODELS` | Model Registry (versioned models, metrics, services) |

## 1. Setup and Connect

```{r}

#| label: setup
# Declare Python deps before reticulate/snowflakeR touch Python, so
# reticulate's ephemeral venv (uv-backed) includes the Snowflake SDKs.
reticulate::py_require(packages = c(
  "snowflake-ml-python>=1.5.0",
  "snowflake-snowpark-python",
  "pandas"
))
library(snowflakeR)
library(RSnowflake)
library(tidymodels)
#library(xgboost)
library(ggplot2)
library(dplyr)
library(scales)
library(reticulate)
source("_config.R")
extra_deps <- c("r-base==4.5.3", "r-xgboost", "r-ranger", "numpy<2.0")
py_require(c("snowflake-connector-python[pandas]", "snowflake-snowpark-python", "pandas")) 
# 1. Instantiate the working Snowpark session via Python SDK using the Workbench credentials 
snowpark <- import("snowflake.snowpark") 
py_session <- snowpark$Session$builder$configs(dict(connection_name = "workbench"))$create() 
cat("tidymodels", as.character(packageVersion("tidymodels")), "loaded\n")
cat("xgboost", as.character(packageVersion("xgboost")), "loaded\n")
```

### Connect and Configure Database Context

```{r}
#| label: connect# 1. Instantiate the working Snowpark session via Python SDK using the Workbench credentials 

# 2. Set active database, schema, and warehouse on the Python session 
# (This populates the session context so snowflakeR's active session check doesn't evaluate to NA) 
py_session$sql("USE WAREHOUSE DEFAULT_WH")$collect() 
py_session$sql("USE DATABASE CREDIT_RISK_ML")$collect() 
py_session$sql("USE SCHEMA MODELS")$collect() 

#snowflakeR connection
conn<-sfr_connect()

tbl_customers <- fqn_source("CREDIT_CUSTOMERS")
tbl_payments  <- fqn_source("CREDIT_PAYMENTS")
tbl_loans     <- fqn_source("CREDIT_LOANS")

#n <- sfr_query(conn, paste("SELECT COUNT(*) AS n FROM", tbl_loans))
#cat(sprintf("Connected. Loan applications: %s\n", format(n$N, big.mark = ",")))
```

### Model Registry

```{r}
#| label: registry
reg <- sfr_model_registry(conn,
  database = CR_DATABASE,
  schema   = CR_REGISTRY_SCHEMA
)
cat(sprintf("Model Registry: %s.%s\n", CR_DATABASE, CR_REGISTRY_SCHEMA))
```

## 2. Feature Store

Define reusable features in the **FEATURES** schema using `dplyr`/`dbplyr`
pipelines. Each Feature View is backed by a Snowflake Dynamic Table that
refreshes automatically.

Features defined here are used identically in training and production scoring —
**define once, use everywhere**.

```{r}

#| label: feature-store-init
fs <- sfr_feature_store(conn,
  database  = CR_DATABASE,
  schema    = CR_FEATURE_SCHEMA,
  warehouse = conn$warehouse,
  create    = TRUE
)
dbi_con <- DBI::dbConnect(
  odbc::snowflake(),
  warehouse = conn$warehouse,
  database  = CR_DATABASE,
  schema    = CR_FEATURE_SCHEMA
)


tbl_customers <- fqn_source("CREDIT_CUSTOMERS")
tbl_payments  <- fqn_source("CREDIT_PAYMENTS")
tbl_loans     <- fqn_source("CREDIT_LOANS")


customers_tbl <- tbl(dbi_con, I(tbl_customers))
payments_tbl  <- tbl(dbi_con, I(tbl_payments))
loans_tbl     <- tbl(dbi_con, I(tbl_loans))
# Assuming 'session' is your active snowpark Session object
# cat("Feature Store and dbplyr table references ready.\n")
# fs
```

### Create Entity

```{r}
#| label: entity
customer_entity <- sfr_create_entity(
  fs, "CUSTOMER", "CUSTOMER_ID",
  desc = "Loan applicant"
)
customer_entity
```

### Feature View 1: Customer Profile

Static demographic features. We build the feature query as a `dbplyr` pipeline,
render to SQL with `dbplyr::sql_render()`, and pass that SQL string to
`sfr_create_feature_view()`.

```{r}
#| label: fv-profile
profile_query <- customers_tbl |>
  select(CUSTOMER_ID, AGE, ANNUAL_INCOME, EMPLOYMENT_LENGTH_YEARS,
         HOME_OWNERSHIP) |>
  mutate(LOG_INCOME = log(pmax(ANNUAL_INCOME, 1)))

profile_sql <- dbplyr::sql_render(profile_query)
cat("dbplyr pipeline rendered to SQL:\n")
cat(as.character(profile_sql), "\n")
```

```{r}
#| label: fv-profile-register
fv_profile <- sfr_create_feature_view(
  fs,
  name      = "CUSTOMER_PROFILE_FV",
  version   = "v1",
  entities  = customer_entity,
  feature_df  = profile_sql,
  desc      = "Customer demographic profile features",
  overwrite = TRUE
)
cat("Registered CUSTOMER_PROFILE_FV (using SQL string)\n")
```

### Feature View 2: Payment Behaviour (Dynamic Table)

Aggregated payment history — captures delinquency patterns. Passes the
`dbplyr` lazy table directly (contrast with the SQL string approach above).

Uses `refresh_freq` to create a **Dynamic Table** that Snowflake refreshes
automatically — use this when source data changes frequently.

```{r}
#| label: fv-payments
payment_query <- payments_tbl |>
  group_by(CUSTOMER_ID) |>
  summarise(
    AVG_DAYS_PAST_DUE  = mean(DAYS_PAST_DUE, na.rm = TRUE),
    MAX_DAYS_PAST_DUE  = max(DAYS_PAST_DUE, na.rm = TRUE),
    PCT_MISSED         = sum(ifelse(PAYMENT_STATUS == "MISSED", 1, 0),
                             na.rm = TRUE) / n(),
    PCT_ON_TIME        = sum(ifelse(PAYMENT_STATUS == "ON_TIME", 1, 0),
                             na.rm = TRUE) / n(),
    AVG_PAYMENT_AMOUNT = mean(PAYMENT_AMOUNT, na.rm = TRUE),
    TOTAL_PAYMENTS     = n(),
    .groups = "drop"
  )

fv_payments <- sfr_create_feature_view(
  fs,
  name         = "PAYMENT_BEHAVIOR_FV",
  version      = "v1",
  entities     = customer_entity,
  feature_df   = payment_query,
  refresh_freq = "1 hour",
  desc         = "Payment behaviour aggregates (dynamic table)",
  overwrite    = TRUE
)
cat("Registered PAYMENT_BEHAVIOR_FV (dynamic table, 1h refresh)\n")
```

### Feature View 3: Credit History

Aggregated lending history. Also passes the `dbplyr` lazy table directly.

```{r}
#| label: fv-credit
credit_query <- loans_tbl |>
  group_by(CUSTOMER_ID) |>
  summarise(
    TOTAL_APPLICATIONS = n(),
    AVG_LOAN_AMOUNT    = mean(LOAN_AMOUNT, na.rm = TRUE),
    MAX_LOAN_AMOUNT    = max(LOAN_AMOUNT, na.rm = TRUE),
    PRIOR_DEFAULTS     = sum(DEFAULTED, na.rm = TRUE),
    AVG_INTEREST_RATE  = mean(INTEREST_RATE, na.rm = TRUE),
    .groups = "drop"
  )

fv_credit <- sfr_create_feature_view(
  fs,
  name      = "CREDIT_HISTORY_FV",
  version   = "v1",
  entities  = customer_entity,
  feature_df  = credit_query,
  desc      = "Credit / lending history aggregates",
  overwrite = TRUE
)
cat("Registered CREDIT_HISTORY_FV\n")
```

```{r}
#| label: list-fvs
cat("Feature Views registered:\n")
sfr_list_feature_views(fs)
```

## 3. Exploratory Data Analysis

```{r}
#| label: eda-data
eda_query <- loans_tbl |>
  left_join(
    customers_tbl |>
      select(CUSTOMER_ID, AGE, ANNUAL_INCOME,
             EMPLOYMENT_LENGTH_YEARS, HOME_OWNERSHIP),
    by = "CUSTOMER_ID"
  )

loans_df <- collect(eda_query)
loans_df$DEFAULTED <- factor(loans_df$DEFAULTED, levels = c(0, 1),
                             labels = c("No Default", "Default"))
cat(sprintf("Joined dataset: %d rows x %d cols\n", nrow(loans_df), ncol(loans_df)))
```

### Default Rate Overview

```{r}
#| label: fig-default-dist
#| fig-cap: "Loan default distribution (~15% default rate)"
ggplot(loans_df, aes(x = DEFAULTED, fill = DEFAULTED)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = comma(after_stat(count))),
            vjust = -0.5, size = 4) +
  scale_fill_manual(values = c("#2ecc71", "#e74c3c")) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Loan Default Distribution",
       subtitle = "~15% default rate -- realistic class imbalance",
       x = NULL, y = "Applications") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

### Income vs Default

```{r}
#| label: fig-income
#| fig-cap: "Lower income correlates with higher default probability"
ggplot(loans_df, aes(x = ANNUAL_INCOME, fill = DEFAULTED)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(labels = dollar) +
  coord_cartesian(xlim = c(0, 250000)) +
  scale_fill_manual(values = c("#2ecc71", "#e74c3c")) +
  labs(title = "Income Distribution by Default Status",
       x = "Annual Income", y = "Density", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "top")
```

### Employment Length vs Default Rate

```{r}
#| label: fig-employment
#| fig-cap: "Longer employment correlates with lower default rate"
emp_summary <- loans_df |>
  mutate(emp_bin = cut(EMPLOYMENT_LENGTH_YEARS,
                       breaks = c(-1, 2, 5, 10, 20, 50),
                       labels = c("0-2yr", "3-5yr", "6-10yr",
                                  "11-20yr", "20+yr"))) |>
  group_by(emp_bin) |>
  summarise(default_rate = mean(DEFAULTED == "Default"),
            n = n(), .groups = "drop")

ggplot(emp_summary, aes(x = emp_bin, y = default_rate)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = sprintf("%.0f%%", default_rate * 100)),
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = percent, expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Default Rate by Employment Length",
       x = "Employment Length", y = "Default Rate") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

## 4. Generate Training Data from Feature Store

Join the label spine (loan applications with default outcome) with all three
Feature Views. The Feature Store handles the joins automatically — the same
features used here will be retrieved at inference time.

```{r}
#| label: training-data
training_data <- sfr_generate_training_data(
  fs,
  spine = paste(
    "SELECT application_id, customer_id, defaulted",
    "FROM", tbl_loans
  ),
  features = list(fv_profile, fv_payments, fv_credit),
  spine_label_cols = "DEFAULTED"
)

tbl_training <- fqn_training("CREDIT_RISK_TRAINING")
sfr_write_table(conn, tbl_training, training_data, overwrite = TRUE)

cat(sprintf("Training data: %d rows x %d cols -> %s\n",
            nrow(training_data), ncol(training_data), tbl_training))
head(as.data.frame(training_data), 5)
```

```{r}
#| label: prep-labels
training_data$DEFAULTED <- factor(training_data$DEFAULTED,
                                  levels = c(1, 0),
                                  labels = c("yes", "no"))

cat("Label distribution:\n")
table(training_data$DEFAULTED)
```

## 5. Model Training with tidymodels

Two models compared with cross-validation:

- **Logistic Regression** — interpretable baseline, standard in credit risk
- **XGBoost** — gradient-boosted trees, typically better predictive performance

### Preprocessing Recipe

```{r}
#| label: recipe
credit_recipe <- recipe(
  DEFAULTED ~ AGE + ANNUAL_INCOME + LOG_INCOME +
    EMPLOYMENT_LENGTH_YEARS + HOME_OWNERSHIP +
    AVG_DAYS_PAST_DUE + MAX_DAYS_PAST_DUE +
    PCT_MISSED + PCT_ON_TIME + AVG_PAYMENT_AMOUNT +
    TOTAL_PAYMENTS + TOTAL_APPLICATIONS + AVG_LOAN_AMOUNT +
    MAX_LOAN_AMOUNT + PRIOR_DEFAULTS + AVG_INTEREST_RATE,
  data = training_data
) |>
  step_dummy(HOME_OWNERSHIP) |>
  step_impute_median(all_numeric_predictors()) |>
  step_zv(all_predictors()) |>
  step_normalize(all_numeric_predictors())

credit_recipe
```

### Define Model Specifications

```{r}
#| label: specs
logistic_spec <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification")

xgb_spec <- boost_tree(
  trees = 500,
  tree_depth = 6,
  learn_rate = 0.05,
  min_n = 10
) |>
  set_engine("xgboost") |>
  set_mode("classification")

cat("Model specifications defined\n")
```

### Create Workflows

```{r}
#| label: workflows
wf_logistic <- workflow() |>
  add_recipe(credit_recipe) |>
  add_model(logistic_spec)

wf_xgb <- workflow() |>
  add_recipe(credit_recipe) |>
  add_model(xgb_spec)

cat("Workflows created\n")
```

### Cross-Validated Evaluation

5-fold stratified CV for honest performance estimates.

```{r}
#| label: cv
set.seed(42)
folds <- vfold_cv(training_data, v = 5, strata = DEFAULTED)
credit_metrics <- metric_set(roc_auc, accuracy, pr_auc)

cat("Evaluating logistic regression (5-fold CV)...\n")
results_lr <- fit_resamples(wf_logistic, folds, metrics = credit_metrics)

cat("Evaluating XGBoost (5-fold CV)...\n")
results_xgb <- fit_resamples(wf_xgb, folds, metrics = credit_metrics)

cat("Cross-validation complete\n")
```

## 6. Model Evaluation

### Compare CV Metrics

```{r}
#| label: cv-comparison
metrics_lr  <- collect_metrics(results_lr) |> mutate(model = "Logistic Regression")
metrics_xgb <- collect_metrics(results_xgb) |> mutate(model = "XGBoost")

comparison <- bind_rows(metrics_lr, metrics_xgb) |>
  select(model, .metric, mean, std_err) |>
  arrange(.metric, desc(mean))

as.data.frame(comparison)
```

```{r}
#| label: fig-cv-compare
#| fig-cap: "Model comparison: 5-fold cross-validation"
ggplot(comparison, aes(x = .metric, y = mean, fill = model)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err),
                position = position_dodge(width = 0.7), width = 0.2) +
  geom_text(aes(label = sprintf("%.3f", mean)),
            position = position_dodge(width = 0.7), vjust = -0.8, size = 3.5) +
  scale_fill_manual(values = c("#3498db", "#e67e22")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Model Comparison: 5-Fold Cross-Validation",
       x = "Metric", y = "Mean Score", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "top")
```

### Train Final Models and ROC Curves

Fit both models on the full training set and compare ROC curves on a holdout.

```{r}
#| label: final-fit
set.seed(42)
split <- initial_split(training_data, prop = 0.8, strata = DEFAULTED)
train_set <- training(split)
test_set  <- testing(split)

fit_lr  <- fit(wf_logistic, data = train_set)
fit_xgb <- fit(wf_xgb, data = train_set)

preds_lr  <- augment(fit_lr, test_set)
preds_xgb <- augment(fit_xgb, test_set)

roc_lr  <- roc_curve(preds_lr, DEFAULTED, .pred_yes) |> mutate(model = "Logistic")
roc_xgb <- roc_curve(preds_xgb, DEFAULTED, .pred_yes) |> mutate(model = "XGBoost")

holdout_lr <- bind_rows(
  roc_auc(preds_lr, DEFAULTED, .pred_yes),
  pr_auc(preds_lr, DEFAULTED, .pred_yes),
  accuracy(preds_lr, DEFAULTED, .pred_class)
) |> mutate(model = "Logistic Regression")

holdout_xgb <- bind_rows(
  roc_auc(preds_xgb, DEFAULTED, .pred_yes),
  pr_auc(preds_xgb, DEFAULTED, .pred_yes),
  accuracy(preds_xgb, DEFAULTED, .pred_class)
) |> mutate(model = "XGBoost")

cat("Holdout metrics:\n")
bind_rows(holdout_lr, holdout_xgb) |>
  select(model, .metric, .estimate) |>
  as.data.frame()
```

```{r}
#| label: fig-roc
#| fig-cap: "ROC curves: Logistic Regression vs XGBoost"
roc_data <- bind_rows(roc_lr, roc_xgb)

auc_lr  <- roc_auc(preds_lr, DEFAULTED, .pred_yes)$.estimate
auc_xgb <- roc_auc(preds_xgb, DEFAULTED, .pred_yes)$.estimate

ggplot(roc_data, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(linewidth = 1.1) +
  geom_abline(linetype = "dashed", color = "grey50") +
  scale_color_manual(values = c("#3498db", "#e67e22"),
    labels = c(sprintf("Logistic (AUC=%.3f)", auc_lr),
               sprintf("XGBoost (AUC=%.3f)", auc_xgb))) +
  labs(title = "ROC Curves: Logistic Regression vs XGBoost",
       x = "False Positive Rate", y = "True Positive Rate", color = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = c(0.7, 0.25))
```

### Select Best Model Locally

```{r}
#| label: best-model
if (auc_xgb >= auc_lr) {
  best_fit  <- fit_xgb
  best_name <- "XGBoost"
  best_auc  <- auc_xgb
} else {
  best_fit  <- fit_lr
  best_name <- "Logistic Regression"
  best_auc  <- auc_lr
}
cat(sprintf("Best model: %s (AUC = %.3f)\n", best_name, best_auc))
```

## 7. Register Models and Select Champion

Register **both** candidate models as versions of `CREDIT_RISK_SCORER` in
the Model Registry, attach holdout metrics, then programmatically select the
champion based on registry-stored metrics.

The registry is the single source of truth for model governance — not local
R variables.

```{r}
#| label: verify-local
sfr_predict_local(best_fit, test_set[1:5, ])
```

```{r}
#| label: register-models
input_cols <- sfr_input_cols(
  training_data,
  exclude = c("DEFAULTED", "CUSTOMER_ID", "APPLICATION_ID")
)
cat(sprintf("Inferred %d input columns from training data\n", length(input_cols)))

extra_deps <- c("r-xgboost", "r-ranger", "numpy<2.0")

mv_lr <- sfr_log_model(
  reg,
  model        = fit_lr,
  model_name   = "CREDIT_RISK_SCORER",
  predict_pkgs = c("tidymodels"),
  input_cols   = input_cols,
  output_cols  = list(prediction = "string"),
  conda_deps   = extra_deps,
  comment      = sprintf("Logistic regression -- AUC %.3f", auc_lr)
)

mv_xgb <- sfr_log_model(
  reg,
  model        = fit_xgb,
  model_name   = "CREDIT_RISK_SCORER",
  predict_pkgs = c("tidymodels", "xgboost"),
  input_cols   = input_cols,
  output_cols  = list(prediction = "string"),
  conda_deps   = extra_deps,
  comment      = sprintf("XGBoost -- AUC %.3f", auc_xgb)
)

cat("Both models registered\n")
```

### Attach Metrics

```{r}
#| label: attach-metrics
for (ver in list(list(mv = mv_lr,  h = holdout_lr),
                 list(mv = mv_xgb, h = holdout_xgb))) {
  for (i in seq_len(nrow(ver$h))) {
    sfr_set_model_metric(reg, "CREDIT_RISK_SCORER",
                         ver$mv$version_name,
                         ver$h$.metric[i], ver$h$.estimate[i])
  }
  cat(sprintf("  %s: attached %d metrics\n",
              ver$mv$version_name, nrow(ver$h)))
}
cat("All metrics stored in Model Registry\n")
```

### Compare Versions from Registry

```{r}
#| label: registry-compare
versions <- sfr_show_model_versions(reg, "CREDIT_RISK_SCORER")

metrics_list <- Filter(Negate(is.null), lapply(versions$name, function(vn) {
  m <- sfr_show_model_metrics(reg, "CREDIT_RISK_SCORER", vn)
  if (length(m) == 0 || is.null(m$roc_auc)) return(NULL)
  data.frame(version = vn,
             roc_auc  = as.numeric(m$roc_auc),
             pr_auc   = as.numeric(m$pr_auc),
             accuracy = as.numeric(m$accuracy))
}))

registry_comparison <- do.call(rbind, metrics_list)
cat("Model versions and metrics from registry:\n")
registry_comparison
```

### Select Champion Version

Pick the version with the highest `roc_auc` and set it as the model's
default version — the version downstream consumers use automatically.

```{r}
#| label: set-champion
champion_idx <- which.max(registry_comparison$roc_auc)
champion_ver <- registry_comparison$version[champion_idx]

sfr_set_default_model_version(reg, "CREDIT_RISK_SCORER", champion_ver)

cat(sprintf("Champion: %s (roc_auc = %.3f, pr_auc = %.3f, accuracy = %.3f)\n",
            champion_ver,
            registry_comparison$roc_auc[champion_idx],
            registry_comparison$pr_auc[champion_idx],
            registry_comparison$accuracy[champion_idx]))
cat("Set as default version in Model Registry\n")
```

## 8. Deploy Champion and Inference

Deploy the champion as a scalable REST service via Snowpark Container Services
with **autocapture** enabled. Autocapture logs every inference request/response
to the model's inference table automatically.

```{r}
#| label: deploy
#| eval: false
sfr_create_compute_pool(conn, "CREDIT_POOL",
                        instance_family = "CPU_X64_M")
sfr_create_image_repo(conn, fqn_model("CREDIT_IMAGES"))

svc_fqn <- paste(CR_DATABASE, CR_REGISTRY_SCHEMA, "CREDIT_RISK_SVC_2", sep = ".")
sfr_execute(conn, paste("DROP SERVICE IF EXISTS", svc_fqn))
```
```{r}
sfr_deploy_model(
  reg,
  model_name   = "CREDIT_RISK_SCORER",
  version_name = champion_ver,
  service_name = "credit_risk_svc_2",
  compute_pool = "CREDIT_POOL",
  image_repo   = fqn_model("CREDIT_IMAGES"),
  force        = TRUE,
  autocapture  = TRUE
)
```

```{r}
#| label: wait-service
#| eval: false
sfr_wait_for_service(reg, "credit_risk_svc_2",
                     timeout_min = 10, poll_sec = 15)
```

### Score New Applications

Retrieve features from the Feature Store for a batch of customers and score
them through the deployed model.

```{r}
#| label: score-rest
#| #| eval: false
scoring_features <- sfr_retrieve_features(
  fs,
  spine = paste(
    "SELECT DISTINCT customer_id FROM", tbl_customers,
    "LIMIT 20"
  ),
  features = list(fv_profile, fv_payments, fv_credit)
)

cat(sprintf("Retrieved %d rows x %d cols from Feature Store\n",
            nrow(scoring_features), ncol(scoring_features)))
```

```{r}
#| label: predict-rest
#| eval: false
predictions <- sfr_predict(
  reg, "CREDIT_RISK_SCORER",
  scoring_features[, names(input_cols)],
  version_name = champion_ver,
  service_name = "credit_risk_svc"
)

results <- cbind(
  scoring_features[, "CUSTOMER_ID", drop = FALSE],
  predictions
)
cat("Predictions for 20 customers:\n")
results
```

### Score from SQL

The SPCS service exposes a `PREDICT` function callable directly from SQL —
accessible to any Snowflake user, BI tool, or scheduled task without R/Python.

```{r}
#| label: predict-sql
#| eval: false
svc_fn   <- paste(CR_DATABASE, CR_REGISTRY_SCHEMA,
                  "CREDIT_RISK_SVC!PREDICT", sep = ".")
col_list <- paste(names(input_cols), collapse = ", ")

sql_score <- paste(
  "SELECT *,", svc_fn, "(", col_list, ") AS prediction",
  "FROM", fqn_training("CREDIT_RISK_TRAINING"),
  "LIMIT 10"
)

cat("SQL prediction query:\n")
cat(sql_score, "\n\n")
sfr_query(conn, sql_score)
```

### Verify Autocapture

Every inference request was logged to the model's inference table (ML
Observability). This feeds drift detection, performance tracking, and
monitoring.

```{r}
#| label: autocapture
#| eval: false
model_fqn <- paste(CR_DATABASE, CR_REGISTRY_SCHEMA,
                   "SFR_CREDIT_RISK_SCORER", sep = ".")

row_count <- 0L
for (attempt in 1:12) {
  row_count <- sfr_query(
    conn,
    sprintf("SELECT COUNT(*) AS N FROM TABLE(INFERENCE_TABLE('%s'))", model_fqn)
  )$N[1]
  if (!is.na(row_count) && row_count > 0) break
  Sys.sleep(5)
}

cat(sprintf("Autocapture rows captured: %d\n", row_count))
stopifnot(
  "No autocapture rows found after 60s" = row_count > 0
)
```

```{r}
#| label: autocapture-detail
#| eval: false
detail_sql <- sprintf("
SELECT
    TIMESTAMP,
    RESOURCE_ATTRIBUTES:\"snow.model.version.name\"::STRING AS MODEL_VERSION,
    RECORD_ATTRIBUTES:\"snow.model_serving.function.name\"::STRING AS FUNCTION_NAME,
    RECORD_ATTRIBUTES:\"snow.model_serving.response.code\"::INT AS RESPONSE_CODE
FROM TABLE(INFERENCE_TABLE('%s'))
ORDER BY TIMESTAMP DESC
LIMIT 10
", model_fqn)

cat("Recent autocapture entries:\n")
sfr_query(conn, detail_sql)
```

## 9. Optional: SQL Pushdown with orbital

If the `orbital` package is installed, the trained tidymodels workflow can be
converted to pure SQL — useful for batch scoring without deploying an SPCS
service.

```{r}
#| label: orbital
#| eval: false
# pak::pak("tidymodels/orbital")
# library(orbital)
# sql_model <- orbital(fit_xgb)
# predict(sql_model, con = dbi_con, table = tbl_training)
```

## 10. Cleanup (Optional)

```{r}
#| label: cleanup
#| eval: false
# Drop SPCS services before deleting the model
# sfr_execute(conn, paste("DROP SERVICE IF EXISTS", svc_fqn))
# sfr_execute(conn, paste("DROP COMPUTE POOL IF EXISTS R_CREDIT_POOL"))

# Drop model
# sfr_execute(conn, paste(
#   "DROP MODEL IF EXISTS",
#   paste(CR_DATABASE, CR_REGISTRY_SCHEMA, "SFR_CREDIT_RISK_SCORER", sep = ".")
# ))

# Full teardown (careful!)
# sfr_execute(conn, paste("DROP DATABASE IF EXISTS", CR_DATABASE))
```

## Summary

| Schema | What we did |
|--------|-------------|
| `RAW_DATA` | Queried source lending data |
| `FEATURES` | Created entity + 3 Feature Views (1 Dynamic Table) |
| `TRAINING` | Generated training data from Feature Store joins |
| `MODELS` | Registered both models, compared metrics, deployed champion |

### Key Advantages Over Traditional Workflow

| Traditional (RStudio + Plumber) | Snowflake (snowflakeR + SPCS) |
|---------------------------------|-------------------------------|
| Data exported to local machine | Data stays in Snowflake |
| Manual feature engineering scripts | Feature Store: define once, use everywhere |
| Model files on shared drives | Model Registry: versioned, governed, with metrics |
| Plumber API on a VM | SPCS: auto-scaled containers, no infra management |
| No inference logging | Autocapture: every prediction logged automatically |
| Manual model comparison | Registry metrics: programmatic champion selection |
